Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-21 paper/sections/metis-language-and-proofs.tex
[ paper ] working on corrections.
[ paper ] putting some order...

\subsection{Input Language.}
\label{ssec:input-language}

@@

The \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, or a \name{hypothesis}.
%existen otros
%----Keywords:    fof cnf thf tff include
%----Punctuation: ( ) , . [ ] :
%----Operators:   ! ? ~ & | <=> => <= <~> ~| ~& * +
%----Predicates:  = != $true $false

%----Top of Page---------------------------------------------------------------
%----CNF formulae (variables implicitly universally quantified)
% <cnf_formula>          ::= <disjunction> | (<disjunction>)
% <disjunction>          ::= <literal> | <disjunction> <vline> <literal>
% <literal>              ::= <fof_atomic_formula> | ~ <fof_atomic_formula> |
%                            <fof_infix_unary>
The formulas include the binary Boolean connectives
(\verb!&!, \verb!|!, \verb!=>!, \verb!<=>!, \verb!~!) and logic constants \verb!$true!
% the boolean type ($true and $false)
and \verb!$false!, for the symbols (∧, ∨, ⇒, ⇔, ¬, ⊤, ⊥), respectively.

@@

This study aimed to address proof-reconstruction for refutation
proofs. Then, we annotated \abbre{CPL} problems using \verb!fof! formulas.
\Metis converts \abbre{FOF} formulas to \abbre{CNF} clauses to look for a
refutation.

\subsection{Output Language.}
\label{ssec:output-language}

The \TSTP language is de facto output standard language
for derivations of ATPs~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006,Sicard-Ramirez2016}.
A \TSTP derivation is a directed acyclic graph, a proof tree,
where each leaf is a formula from the \TPTP input. A node is a formula
inferred from the parent formulas. The root is the final derived formula,
such a derivation is a list of annotated formulas with the form.

@@

The \name{rule} stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning
step, and the list of parents nodes.

@@

In the script above, \strip is for example the name of the inference.
It has no arguments and derives from one parent node with goal name.
\name{goal}.

@@

\subsection{Inference Rules.}
\label{ssec:metis-inferences-rules}

@@

Finally, \emph{axiom} rule appears in \TSTP derivations by using one of the following rules: \canonicalize, \clausify, \conjunct, or \simplify.
Then, the last rule remains is \emph{resolve}, it
appears by the binary rule, \resolve, the version in \Metis of the resolution
theorem. For a more detailed description of this rules see Section~\ref{sec:proof-reconstruction}.

\subsection{Proofs.}
\label{ssec:metis-proofs}

@@

\begin{figure}
\centering
  \begin{bprooftree}\tt
    \AxiomC{}
    \RightLabel{negate}
    \UnaryInfC{$\neg p$}
    \RightLabel{strip}
    \UnaryInfC{$\neg p$}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$p$}
    \RightLabel{canonicalize}
    \UnaryInfC{$p$}
    \RightLabel{simplify}
    \BinaryInfC{$⊥$}
    \RightLabel{canonicalize}
    \UnaryInfC{$⊥$}
  \end{bprooftree}
  \caption{\Metis' proof tree for $p \vdash p$. Derivation in Fig.~\ref{fig:metis-proof-tstp}}
  \label{fig:metis-example}
\end{figure}

The output of a proof delivered by \Metis encodes a natural
deduction proof. The deduction system uses six inference rules
\cite{hurd2003first} listed in Fig.~\ref{fig:metis-inferences}.
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (e.g.,
\emph{falsium} in the last conclusion of the \TSTP derivation in
Fig.~\ref{fig:metis-example}).

These proofs are directed acyclic graphs, trees of refutations. Each
node stands for an application of an inference rule and the leaves
in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule (e.g., \canonicalize in
Fig.~\ref{fig:metis-example}). Each edge links a premise with one
conclusion. The proof graphs have at their root the conclusion
$⊥$, since \Metis delivers proofs by refutation.

>>>>

\subsection{Input language}
\label{ssec:input-language}

@@

The \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, or a \name{hypothesis}.
The formulas include the binary connectives
(\verb!&!, \verb!|!, \verb!=>!, \verb!<=>!, \verb!~!)
and the constants \verb!$true!, and \verb!$false!.

@@

\subsection{Output language}
\label{ssec:output-language}

The \TSTP language is an output language for derivations of
ATPs~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006,Sicard-Ramirez2016}.
A \TSTP derivation is a directed acyclic graph, a proof tree,
where each leaf is a formula from the \TPTP input. A node is a formula
inferred from the parent formulas. The root is the final derived formula,
such a derivation is a list of annotated formulas with the following form.

@@

The \name{rule} in the line above stands for the inference name;
the other fields are supporting arguments or useful information to
apply the reasoning step, and list the parents nodes.

@@

In the script above, \strip is for example the name of the inference.
It has no arguments and derives from one parent node named goal. The
result of this inference when it applies to the \verb!goal! formula is
\verb!p!.

@@

\subsection{Inference rules}
\label{ssec:metis-inferences-rules}

@@

% Finally, \emph{axiom} rule appears in \TSTP derivations by using one of the following rules: \canonicalize, \clausify, \conjunct, or \simplify.
% Then, the last rule remains is \emph{resolve}, it
% appears by the binary rule, \resolve, the version in \Metis of the resolution
% theorem. For a more detailed description of this rules see Section~\ref{sec:proof-reconstruction}.

\subsection{Proofs}
\label{ssec:metis-proofs}

@@

The output of a proof delivered by \Metis encodes a natural
deduction proof (see a complete proof in Fig.~\ref{fig:metis-proof-tstp}.
The deduction system uses six inference rules \cite{hurd2003first}.
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (e.g.,
\emph{falsium} in the last conclusion in the \TSTP derivation
Fig.~\ref{fig:metis-proof-tstp}).

These proofs are directed acyclic graphs, trees of refutations
(see the proof tree of the \TSTP derivation showed above in
Fig.~\ref{fig:metis-proof-tstp}).
Each node stands for an application of an inference rule and the leaves
in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule (e.g., \canonicalize).
Each edge links a premise with one conclusion.
The proof graphs have at their root the conclusion
$⊥$, since \Metis delivers proofs by refutation.

\begin{figure}[!ht]
\centering
  \begin{bprooftree}\tt
    \AxiomC{}
    \RightLabel{negate}
    \UnaryInfC{$\neg p$}
    \RightLabel{strip}
    \UnaryInfC{$\neg p$}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$p$}
    \RightLabel{canonicalize}
    \UnaryInfC{$p$}
    \RightLabel{simplify}
    \BinaryInfC{$⊥$}
    \RightLabel{canonicalize}
    \UnaryInfC{$⊥$}
  \end{bprooftree}
  \caption{Proof tree for the \Metis derivation in
  Fig.~\ref{fig:metis-proof-tstp}}.
  \label{fig:metis-example}
\end{figure}

### 2017-10-21 paper/sections/proof-reconstruction.tex
[ paper ] putting some order...

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the formalizations of \Metis inferences.

In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user, checking for the error in \Metis, \Athena,
or in \Agda.

@@

Therefore, for a general recursive function $\rm{f}: \Prop \to \Prop$
with termination problems, we state a new function
$\rm{f}^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must be
done by using induction on the \Nat type structure.

\subsubsection*{An  example.}
Let us consider the following example to exhibit the bounded technique
mentioned above for the \fuh function presented in the version
proposed of the \Metis \strip inference.

@@

Therefore, we define the function
$uh_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:uh-structured}).

>>>>

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the proposed formalizations for the
propositional logic and the \Metis inferences.
In that case \Agda outputs an interface file.

Otherwise, when type-checking fails, the failure must be
investigated by the user looking at the error in the \TSTP derivation by \Metis,
in the translation by \Athena,in the \Agda formalisations mentioned above or
even in the type-checker, \Agda..

@@

Therefore, for a general recursive function $\rm{f}: \Prop \to \Prop$
with termination problems, we formulate a new function
$\rm{f}^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must be
done by using induction on the \Nat type structure.

\subsubsection*{An  example.}
Let us consider the following example to exhibit the bounded technique
mentioned above for the \fuh function presented in the proposed version
for the \Metis \strip inference.

@@

Therefore, we define the function
$\fuh_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:uh-structured}).
